Weekday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population

### 2023-12-11 manuscript/paper.Rmd
upd

```{r source-file, echo=FALSE, include=FALSE}
# source file 
# source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
```

@@

<!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized -->
<!-- contributor to systolic blood pressure (SBP), with individuals -->
<!-- exhibiting higher BMI values demonstrating a propensity for elevated -->
<!-- SBP. Gender also holds a discernible association with SBP dynamics. -->
<!-- Empirical evidence in CITE demonstrates that women typically manifest -->
<!-- lower SBP than men until the onset of menopause, after which women -->
<!-- exhibit higher SBP levels compared to men. Additionally, heightened -->
<!-- levels of cholesterol are closely linked to vascular damage. The -->
<!-- accumulation of cholesterol deposits and calcium within arteries -->
<!-- precipitates arterial stiffening, necessitating augmented cardiac -->
<!-- efforts to circulate blood, potentially culminating in elevated blood -->
<!-- pressure. This study endeavors to examine the interrelationships between -->
<!-- age and cholesterol levels, emphasizing consideration of potential -->
<!-- confounding variables. Table 1 comprehensively delineates the array of -->
<!-- variables under scrutiny, ensuring a comprehensive analysis of their -->
<!-- collective impact -->

@@

Sleep duration on workdays was evaluated by the questionnaire with the
following questions: "Number of hours usually sleep on weekdays or
workdays". We then categorized this variable into three groups, that is, \< 6
h, 6--8 h, ≥ 8 h respectively, and used < 6 h as the reference
group in our analysis. We also explored the association between total cholestrol (in mmol/L) level on systolic blood pressure including potential effect modification by other factors.

@@

\begin{table}[H]

\caption{\label{tab:tabone}Variable descriptions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Name} & \textbf{Description}\\
\midrule
\textbf{SEQN} & \textbf{sequence number} & Respondent number\\
\textbf{SDMVPSU} & \textbf{psu} & Masked variance unit pseudo-PSU variable for variance estimation\\
\textbf{WTINT2YR} & \textbf{weights} & Full sample 2-year interview weights\\
\textbf{SDMVSTRA} & \textbf{strata} & Masked variance unit pseudo-stratum variable for variance estimation\\
\textbf{RIAGENDR} & \textbf{gender} & Respondent's number\\
\addlinespace
\textbf{RIDAGEYR} & \textbf{age (yrs)} & Respondent's age in years\\
\textbf{DMDMARTL} & \textbf{marital status} & Marital status\\
\textbf{INDFMIN2} & \textbf{income category} & Total family income (reported as a range value in dollars)\\
\textbf{RIDRETH3} & \textbf{race} & Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category\\
\textbf{DMDHHSZA} & \textbf{children <5} & Number of children aged 5 years or younger in the household\\
\addlinespace
\textbf{DMDEDUC2} & \textbf{education level} & What is the highest grade or level of school\{you have/SP has\} completed or the highest degree \{you have/s/he has\} received?\\
\textbf{DMDCITZN} & \textbf{citizenship status} & \{Are you/Is SP\} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv\\
\textbf{SLD012} & \textbf{sleep} & Number of hours usually sleep on weekdays or workdays\\
\textbf{BMXBMI} & \textbf{bmi} & Body mass index\\
\textbf{ALQ121} & \textbf{alcohol use} & In the past 12 months, how often did you drink any type of alcoholic beverage?\\
\addlinespace
\textbf{LBDHDDSI} & \textbf{hdl} & High density lipoprotein\\
\textbf{LBDSALSI} & \textbf{albumin} & Albumin (g/L)\\
\textbf{DIQ010} & \textbf{diabetes} & Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?\\
\textbf{BPQ020} & \textbf{hypertension} & \{Have you/Has SP\} ever been told by a doctor or other health professional that \{you/s/he\} had hypertension, also called high blood pressure?\\
\textbf{URXCRS} & \textbf{creatinine} & Creatinine, urine (umol/L)\\
\addlinespace
\textbf{SLQ040} & \textbf{snort} & In the past 12 months, how often did \{you/SP\} snort, gasp, or stop breathing while \{you were/s/he was\} asleep?\\
\textbf{LBDTCSI} & \textbf{total cholestrol levels} & Total cholesterol (mmol/L)\\
\textbf{LBXHGB} & \textbf{hemoglobin} & Hemoglobin (g/dL)\\
\textbf{LBXSASSI} & \textbf{AST} & Aspertate aminotransferase\\
\textbf{SMQ040} & \textbf{smoke} & Do you now smoke cigarettes?\\
\addlinespace
\textbf{BPXDI} & \textbf{blood pressure} & Systolic and diastolic blood pressure taken as the average of the first three measurements\\
\bottomrule
\end{tabular}
\end{table}

```{r tabone, echo=FALSE, eval=FALSE}
tab.one <- vars %>% 
  kable(format = "latex", caption = "Variable descriptions", booktabs = TRUE) %>% 
  kable_styling(latex_options = c("HOLD_position")) %>%
  row_spec(c(0), bold = TRUE) %>%
  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>%
  column_spec(3, width = "8cm")
tab.one
```

@@

\begin{table}

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & df & AIC & BIC & MSE & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 41 & 15888.38 & 16117.55 & 173.69 & 0.36 & 0.34\\
Model 7 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}

```{r tabmodel, eval=FALSE, echo=FALSE}
## Model performance
tabmodel <- model.checks %>% 
  kable(format = "latex", caption = "Summary model performance measures.", digits = 2, booktabs = TRUE, col.names = c("df", "AIC", "BIC", "MSE", "R^2", "Adjusted R2"), escape = FALSE) %>% 
  add_footnote(label = c("df = degrees of freedom", "AIC = Akaike Information Criteria", "BIC = Bayesian Information Criteria", "MSE = Mean Squared Error"))
tabmodel
```

>>>>

```{r source-file, echo=FALSE, include=FALSE}
# source file 
source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")

#source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
```

@@

<!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized -->

<!-- contributor to systolic blood pressure (SBP), with individuals -->

<!-- exhibiting higher BMI values demonstrating a propensity for elevated -->

<!-- SBP. Gender also holds a discernible association with SBP dynamics. -->

<!-- Empirical evidence in CITE demonstrates that women typically manifest -->

<!-- lower SBP than men until the onset of menopause, after which women -->

<!-- exhibit higher SBP levels compared to men. Additionally, heightened -->

<!-- levels of cholesterol are closely linked to vascular damage. The -->

<!-- accumulation of cholesterol deposits and calcium within arteries -->

<!-- precipitates arterial stiffening, necessitating augmented cardiac -->

<!-- efforts to circulate blood, potentially culminating in elevated blood -->

<!-- pressure. This study endeavors to examine the interrelationships between -->

<!-- age and cholesterol levels, emphasizing consideration of potential -->

<!-- confounding variables. Table 1 comprehensively delineates the array of -->

<!-- variables under scrutiny, ensuring a comprehensive analysis of their -->

<!-- collective impact -->

@@

Sleep duration on workdays was evaluated by the questionnaire with the
following questions: "Number of hours usually sleep on weekdays or
workdays". We then categorized this variable into three groups, that is,
\< 6 h, 6--8 h, ≥ 8 h respectively, and used \< 6 h as the reference
group in our analysis. We also explored the association between total
cholestrol (in mmol/L) level on systolic blood pressure including
potential effect modification by other factors.

@@

```{=tex}
\begin{table}[H]

\caption{\label{tab:tabone}Variable descriptions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Name} & \textbf{Description}\\
\midrule
\textbf{SEQN} & \textbf{sequence number} & Respondent number\\
\textbf{SDMVPSU} & \textbf{psu} & Masked variance unit pseudo-PSU variable for variance estimation\\
\textbf{WTINT2YR} & \textbf{weights} & Full sample 2-year interview weights\\
\textbf{SDMVSTRA} & \textbf{strata} & Masked variance unit pseudo-stratum variable for variance estimation\\
\textbf{RIAGENDR} & \textbf{gender} & Respondent's number\\
\addlinespace
\textbf{RIDAGEYR} & \textbf{age (yrs)} & Respondent's age in years\\
\textbf{DMDMARTL} & \textbf{marital status} & Marital status\\
\textbf{INDFMIN2} & \textbf{income category} & Total family income (reported as a range value in dollars)\\
\textbf{RIDRETH3} & \textbf{race} & Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category\\
\textbf{DMDHHSZA} & \textbf{children <5} & Number of children aged 5 years or younger in the household\\
\addlinespace
\textbf{DMDEDUC2} & \textbf{education level} & What is the highest grade or level of school\{you have/SP has\} completed or the highest degree \{you have/s/he has\} received?\\
\textbf{DMDCITZN} & \textbf{citizenship status} & \{Are you/Is SP\} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv\\
\textbf{SLD012} & \textbf{sleep} & Number of hours usually sleep on weekdays or workdays\\
\textbf{BMXBMI} & \textbf{bmi} & Body mass index\\
\textbf{ALQ121} & \textbf{alcohol use} & In the past 12 months, how often did you drink any type of alcoholic beverage?\\
\addlinespace
\textbf{LBDHDDSI} & \textbf{hdl} & High density lipoprotein\\
\textbf{LBDSALSI} & \textbf{albumin} & Albumin (g/L)\\
\textbf{DIQ010} & \textbf{diabetes} & Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?\\
\textbf{BPQ020} & \textbf{hypertension} & \{Have you/Has SP\} ever been told by a doctor or other health professional that \{you/s/he\} had hypertension, also called high blood pressure?\\
\textbf{URXCRS} & \textbf{creatinine} & Creatinine, urine (umol/L)\\
\addlinespace
\textbf{SLQ040} & \textbf{snort} & In the past 12 months, how often did \{you/SP\} snort, gasp, or stop breathing while \{you were/s/he was\} asleep?\\
\textbf{LBDTCSI} & \textbf{total cholestrol levels} & Total cholesterol (mmol/L)\\
\textbf{LBXHGB} & \textbf{hemoglobin} & Hemoglobin (g/dL)\\
\textbf{LBXSASSI} & \textbf{AST} & Aspertate aminotransferase\\
\textbf{SMQ040} & \textbf{smoke} & Do you now smoke cigarettes?\\
\addlinespace
\textbf{BPXDI} & \textbf{blood pressure} & Systolic and diastolic blood pressure taken as the average of the first three measurements\\
\bottomrule
\end{tabular}
\end{table}
```
```{r tabone, echo=FALSE, eval=FALSE}
tab.one <- vars %>% 
  kable(format = "latex", caption = "Variable descriptions", booktabs = TRUE) %>% 
  kable_styling(latex_options = c("HOLD_position")) %>%
  row_spec(c(0), bold = TRUE) %>%
  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>%
  column_spec(3, width = "8cm")
tab.one
```

@@

```{=tex}
\begin{table}

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & df & AIC & BIC & MSE & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 41 & 15888.38 & 16117.55 & 173.69 & 0.36 & 0.34\\
Model 7 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}
```
```{r tabmodel, eval=FALSE, echo=FALSE}
## Model performance
tabmodel <- model.checks %>% 
  kable(format = "latex", caption = "Summary model performance measures.", digits = 2, booktabs = TRUE, col.names = c("df", "AIC", "BIC", "MSE", "R^2", "Adjusted R2"), escape = FALSE) %>% 
  add_footnote(label = c("df = degrees of freedom", "AIC = Akaike Information Criteria", "BIC = Bayesian Information Criteria", "MSE = Mean Squared Error"))
tabmodel
```
